\documentclass{article}
\usepackage{times}
\usepackage{fullpage}

\newcommand{\head}{\subsection*}
\setlength{\parindent}{0pt}

\begin{document}
%Test =)
%_______________________________________________________________________________
% Heading section
%_______________________________________________________________________________
\begin{center}
\rule{\textwidth}{1.5pt} \\ \rule[10pt]{\textwidth}{1pt}\\
MSIT-SE-M-01\hfill Models of Software Systems\\[3ex]
{\Large\bf Project 2: Z Modeling}\\[3ex]
Innopolis University \hfill {\bf Due: October 28, 2015} \rule{\textwidth}{1pt}
\\\rule[9.5pt]{\textwidth}{1.5pt}
\end{center}

The purpose of this project is to give you experience in modeling a
realistic system as a Z specification. The example that we will use
is the Infusion Pump. A general description of an Infusion Pump can
be found in the General Project Documents folder on the class
blackboard.

\bigskip You should carry out this project in your assigned team.
Make sure that everyone in the group contributes to the overall
effort. Each team should submit a single write-up of the project,
due at the beginning of class on the project due date. We have
posted a template for a group project write-up under the
\LaTeX~section of the course web site.

%___________________________________________________________________________________________________
\head{Task 1 (90 points):}
%___________________________________________________________________________________________________

From a certain level of abstraction, a machine with multiple,
concurrent threads may be considered to be a machine with a single
thread, particularly if we do not permit events to occur
simultaneously. From this point of view, it is possible to treat an
N-channel infusion pump as a sequential system where events may
occur in any one channel, but at different times. Moreover, the
specification of events' timing is a non-trivial task and we can
abstract away when the events must occur.

\bigskip Your first task is to develop a specification of a
3-channel pump. Use the description of the model from the Blackboard
site to define the requirements for the pump. Your model should show
how the pump operates  in the presence of failures and, also,
consider the behavior in the presence or absence of electric power.
As a hint, consider separating the model of a channel from the model
of a pump that houses 3-channels. You should use promotion (just as
in the Library Problem) to extend a model of a channel to the model
of the pump.

\bigskip Be sure to document your specification with appropriate explanatory prose

\bigskip Effective use of the schema calculus will be one of the evaluation criteria for your specification.

\bigskip Be sure to typecheck your specification using fuzz,
Z-Eves, or CZT.

\bigskip Answer the following questions in your project write-up:
\begin{enumerate}
    \item Describe any decisions that your team made in resolving
    ambiguities in the English description. Also describe
    alternatives you considered and rejected.
    \item Would your model be more complicated if there were more
    channels?
    \item What happens when the pump is out of liquid?
    \item Is it possible to dispense medicine without first
    setting a rate?
    \item Is it possible to start the pump when the key is locked?
    \item What happens when the the pump is out of power?
    \item Are there any extensions in your model not defined in
    the written description of the pump?
\end{enumerate}

\noindent The full specification should be attached to your project
write-up. Hand in a hard copy and email a soft copy to the course
teaching assistants before class on the due date.

%___________________________________________________________________________________________________
\head{Task 2 (10 Points):}
%___________________________________________________________________________________________________

For the second part of your project you should write a short essay
(less than 1 page) that speculates on a more-complete Z model of the infusion pump. In particular, consider how you might use the Z schema calculus to introduce start- and end-times and duration into your model.
Additionally, reflect on the experience of developing a model in a
group; concentrate on how the formal model helped or hindered you in
understanding the infusion pump.


\end{document}
